---
title: "Lab 1: Introduction to R"
output: html_notebook
---


> #### Objects

* Objects should have names
* Object Types: vector, matrix … etc.
* Object Attributes
    + mode: numeric, character, boolean
    + length: number of elements in object
* Object Values
    + assign a value
    + create a blank object

> #### Naming Convention 

* must start with a letter (A-Z or a-z)
* can contain letters, digits (0-9), and/or periods “.” ex: Var1.1
* case-sensitive
    + mydata different from MyData
* do not use underscore “_”

> #### Assignment 

**<-** used to indicate assignment

```{r}
x <- 1
y <- 3
z <- 4
x*y*z
```

**Note:** Type determined automatically when variable is created with **"<-"** operator

> #### Built-In Functions

* actions can be performed on objects using functions
* have arguments and options
* provide a result
* parentheses () are used to specify that a function is being called

> Example Functions

```{r}
rep(1,10)
```

```{r}
seq(2,6)
```

```{r}
x <- c(2,0,0,4)
```

```{r}
x * 4 
```

```{r}
sqrt(x)
```


> #### Objects | Vectors 

* A series of numbers 
* Created with:
    + c() to concatenate elements or sub-vectors 
    + rep() to repeat elements or patterns 
    + seq() or m:n to generate sequences 

> Example

```{r}
X <- c(2,0,0,9)
Y <- seq(2,5) #sequence of integers between 2 & 5
Z <- rep(1,4) #repeat the number 1, 4 times 
X+Y+Z 
```


```{r}
x <- c (2,0,0,4)
x [1] # Select the first element, equivalent to x[c(1)] 
```

```{r}
x [-1] # Exclude the first element
```

```{r}
 x [1] <- 3 ; x
```

```{r}
x [-1] = 5 ; x
```

```{r}
x<5
```

```{r}
x [x<5] = 2 #Edits elements meeting condition
```


> #### Objects | Data Frames 

* A group or collection of Vectors
* Most of the time, when data is loaded, it will be organized in a data frame 

> Example

```{r}
DF <- data.frame (h=c(150,160), w=c(65,72))
DF
```
```{r}
DF[1]
```
```{r}
DF[2]
```
```{r}
DF[1,]
```
```{r}
DF[2,]
```

> #### Objects | Lists

* Collections of related variables 
* Similar to records in C 
* Created with list function 

```{r}
point <- list(x = 1, y = 1) 
```
* Access to components follows similar rules as for data frames, the following all retrieve x: 

```{r}
point$x 
point["x"] 
point[1]
point[-2] 
```

> #### repeat

```{r}
x <- 1
repeat 
{ x <- x+1; 
if (x==50) 
break;}
x
```
> ##### while

* while (expr_1) expr_2 While expr_1 is false, repeatedly evaluate expr_2 
* break and next statements can be used within the loop

```{r}
while (x!=60)
{ x <- x+1;}

x
```

> ### Thank You & Good Luck 